Coursera Reproducible Research Assignment 1 - 7/19/2014
========================================================

## Set Directory and Read in Data

```{r reading in data}
setwd("C:/Cousera/Course5_Reproducible Research")
ActivityData <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
## Make a histogram of the total number of steps taken each day
## Calculate and report the mean and median total number of steps taken per day

## Summarize to each day by summing steps per day - omitting NA and Histogram

```{r histogram for Steps}
StepByDay <- aggregate(steps ~ date, data=ActivityData, FUN=sum,na.action = na.omit)
library (ggplot2)
ggplot (StepByDay,aes(date)) 	+ geom_histogram(aes(weight=steps, fill =..count..),binwidth=.75) + xlab("Date") + theme(axis.text.x=element_text(angle = 90)) + ylab("Steps") + ggtitle("Total Steps Taken Per Day")
```

## Mean and Median per day

```{r mean and median calc}
MeanStepByDay <- aggregate(steps ~ date, data=ActivityData, FUN=mean,na.action = na.omit)
MedStepByDay <- aggregate(steps ~ date, data=ActivityData, FUN=median,na.action = na.omit)

MeanStepByDay
MedStepByDay 
```

## What is the average daily activity pattern?

```{r average daily activity}
ActivityByDay <- aggregate(steps~interval, data=ActivityData, FUN=mean,na.action = na.omit)
ggplot (ActivityByDay ,aes(interval,steps)) + geom_line() + xlab("Interval") + theme(axis.text.x=element_text(angle = 90)) + ylab("Ave Steps") + ggtitle("Average Daily Activity")
```

## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

```{r max steps}
MaxInterval <- max(ActivityByDay$interval)
MaxRow <- which.max(ActivityByDay$steps)
MaxStep <- ActivityByDay$steps[MaxRow]

MaxStep
MaxInterval
```

## Imputing missing values - Calculate and report the total number of missing values in the dataset
Using Not complete cases
```{r missing values}
Missing <- sum(!complete.cases(ActivityData))
Missing
```

## Create a new dataset that is equal to the original dataset but with the missing data filled in. 
## My strategy is to replace the NA with a total average
First, get the overall mean steps, then find the Row that has NA, replace the NA with the Mean
```{r replace NA for Mean}
MeanStep <- mean(ActivityByDay$steps)
ActivityImpute <- ActivityData
NARow <- is.na(ActivityData$steps)
ActivityImpute[NARow,1] <-  MeanStep 
```

## Summarize to each day by summing steps per day - with new NA

```{r new Summary stpe and histogram}
StepByDayImp <- aggregate(steps ~ date, data=ActivityImpute, FUN=sum)
library (ggplot2)
ggplot (StepByDayImp ,aes(date)) + geom_histogram(aes(weight=steps, fill =..count..),binwidth=.75) + xlab("Date") + theme(axis.text.x=element_text(angle = 90)) + ylab("Steps") + ggtitle("Total Steps Taken Per Day")
```

# Mean and Median per day

```{r mean and median for new no NA file}
MeanStepByDayImp <- aggregate(steps ~ date, data=ActivityImpute, FUN=mean,na.action = na.omit)
MedStepByDayImp <- aggregate(steps ~ date, data=ActivityImpute, FUN=median,na.action = na.omit)

MeanStepByDayImp
MedStepByDayImp 
```

## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels – “weekday” and 
“weekend” indicating whether a given date is a weekday or weekend day.

## add the weekday into dataset
Using Find and Replace - found on web hack - then column bind

```{r weekend comparison}
Wkend <- weekdays(as.Date(ActivityImpute$date), abbreviate = TRUE)
Wkend <- sub("^Mon|^Tue|^Wed|^Thu|^Fri", "weekday", Wkend ,  perl = TRUE)
Wkend <- sub("^Sat|^Sun", "weekend", Wkend ,  perl = TRUE)

ActivityImputeW <- cbind(ActivityImpute,Wkend)
```

## What is the average daily activity pattern?
Same graph as above but with facet - remember the + Wkend column for the added aggregation
```{r facet line graph}
ActivityImpByDay <- aggregate(steps~interval + Wkend, data=ActivityImpute , FUN=mean)
ggplot (ActivityImpByDay ,aes(interval,steps)) + geom_line() + facet_grid(Wkend ~ .) + xlab("Interval") + theme(axis.text.x=element_text(angle = 90)) + ylab("Ave Steps") + ggtitle("Average Daily Activity") 
```
